A Phenomenological Model for the Intra-Cluster Medium that matches X-ray and Sunyaev-Zel'dovich observations

### 2013-09-24 paper1.tex
abstract rewritten

%MNRAS
\title[A Phenomenological Model for the Intra-Cluster Medium]{A Phenomenological Model for the Intra-Cluster Medium that matches X-ray and Sunyaev-Zel'dovich observations}
\author[F. Zandanel, C. Pfrommer and F. Prada]{
Fabio Zandanel$^{1,2 *}$, Christoph Pfrommer$^{3 \dagger}$ and Francisco Prada$^{4,5,1 \S}$\\
$^{1}$Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (CSIC), Glorieta de la Astronom\'{\i}a, E-18080 Granada, Spain\\
$^{2}$now at GRAPPA Institute, University of Amsterdam, Science Park 904, 1098XH Amsterdam, Netherlands\\
$^{3}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany\\
$^{4}$Campus of International Excellence UAM+CSIC, Cantoblanco, E-28049 Madrid, Spain\\
$^{5}$Instituto de F\'{\i}sica Te\'orica, (UAM/CSIC), Universidad Aut\'onoma de Madrid, Cantoblanco, E-28049 Madrid, Spain}
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf We construct a phenomenological gas density model for the intra-cluster medium
  based on X-ray inferred densities for cool-core and non cool-core clusters, and a 
  cluster-mass-dependent gas fraction. We then build a mock galaxy cluster 
  catalog from the large MultiDark $N$-body $\Lambda$CDM simulation by assigning
  such a (mass-dependent) gas density profile to each dark matter halo in our sample, 
  which is sorted into cool-core and non cool-core clusters according to a dynamical 
  disturbance parameter. Our gas density model permits to match the observed 
  Sunyaev-Zel'dovich and X-ray scaling relations and luminosity function. Therefore,
  it can be reliable applied to make predictions for present and future instruments. 
  We apply our catalogs to calculate non-thermal radio and gamma-ray emission of galaxy 
  clusters in a companion paper. Eventually, we make our cosmologically complete multi-frequency 
  mock catalogs for the (non-)thermal cluster emission at different redshifts publicly and freely 
  available on-line through the MultiDark database.}
\end{abstract}

@@

\begingroup
\let\thefootnote\relax\footnotetext{* f.zandanel@uva.nl}
\let\thefootnote\relax\footnotetext{$\dagger$ christoph.pfrommer@h-its.org}
\let\thefootnote\relax\footnotetext{$\S$ fprada@iaa.es\\}
\endgroup

{\bf The study of the (non-)thermal emission in galaxy clusters provides
important information on structure formation process and on 
fundamental characteristics of our Universe. This field is at a 
crucial point of its development thanks to the very low sensitivity of present and future-generation 
instruments such as LOFAR\footnote{www.lofar.org}, \emph{e}ROSITA\footnote{www.mpe.mpg.de/eROSITA}, 
ASTRO-H\footnote{astro-h.isas.jaxa.jp} and CTA\footnote{www.cta-observatory.org}, to name but a few. 
Therefore, it is important to provide prospects for future surveys and
to understand in which way they can improve our current knowledge.
To this end, a complete cosmological mock sample of galaxy
clusters is needed. This can be realized either using analytical 
expressions for the cluster mass function (e.g., \citealp{2001MNRAS.321..372J}) or cosmological cluster 
simulations. In this work, we make use of the recent MultiDark $N$-body
simulation (\citealp{2011arXiv1104.5130P}; www.multidark.org).

@@

In next Section~\ref{sec:2}, we introduce the MultiDark simulation
along with our final cluster sample. We assign to each of our clusters  
the phenomenologically constructed gas density profile in 
Section~\ref{sec:3}. In Sections~\ref{sec:4} and 
\ref{sec:5}, we show that this approach can successfully reproduce 
the known X-ray and SZ cluster characteristics. We present the
resulting mock cluster catalogs in Section~\ref{sec:6}, and eventually 
summarize our conclusions in Section~\ref{sec:7}.
In this work, the cluster mass $M_{\Delta}$ and radius $R_{\Delta}$ are defined 
with respect to a density that is $\Delta=200$ or $\Delta=500$ times the \emph{critical} 
density of the Universe. We adopt density parameters of $\Omega_{\rmn{m}}=0.3$, 
$\Omega_{\Lambda}=0.7$ and today's Hubble constant of 
$H_0 = 100 \times h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.
}

>>>>

%MNRAS
\title[A Phenomenological Model for the Intra-Cluster Medium]{A Phenomenological Model for the Intra-Cluster Medium that matches X-ray and Sunyaev-Zel'dovich observations}
\author[F. Zandanel, C. Pfrommer and F. Prada]{
Fabio Zandanel$^{1,2 *}$, Christoph Pfrommer$^{3}$ and Francisco Prada$^{4,5,1}$\\
$^{1}$Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (CSIC), Glorieta de la Astronom\'{\i}a, E-18080 Granada, Spain\\
$^{2}$now at GRAPPA Institute, University of Amsterdam, Science Park 904, 1098XH Amsterdam, Netherlands\\
$^{3}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany\\
$^{4}$Campus of International Excellence UAM+CSIC, Cantoblanco, E-28049 Madrid, Spain\\
$^{5}$Instituto de F\'{\i}sica Te\'orica, (UAM/CSIC), Universidad Aut\'onoma de Madrid, Cantoblanco, E-28049 Madrid, Spain}
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Cosmological hydrodynamical simulations of galaxy clusters are still
  challenged to obtain a model for the intracluster medium that matches all
  aspects of current X-ray and Sunyaev-Zel'dovich observations. To facilitate
  such comparisons with future simulations and to enable realistic cluster
  population studies for modeling e.g., non-thermal emission processes, we
  construct a phenomenological model for the intracluster medium that is based
  on a representative sample of X-ray clusters. We create a mock galaxy cluster
  catalog by assigning our gas density model to each cluster halo of the large
  collisionless $N$-body simulation {\em MultiDark}, which employs a $\Lambda$
  cold dark matter cosmology. Our clusters are identified with cool-core and non
  cool-core clusters according to a dynamical disturbance parameter. We
  demonstrate that our gas model matches the various observed Sunyaev-Zel'dovich
  and X-ray scaling relations as well as the X-ray luminosity function, thus
  enabling a reliable mock catalog for present and future instruments.  In a
  companion paper, we apply our catalogs to calculate non-thermal radio and
  gamma-ray emission of galaxy clusters. We make our cosmologically complete
  multi-frequency mock catalogs for the (non-)thermal cluster emission at
  different redshifts publicly and freely available online through the MultiDark
  database.
\end{abstract}

@@

\begingroup
\let\thefootnote\relax\footnotetext{* f.zandanel@uva.nl}
\endgroup

The study of the (non-)thermal emission in galaxy clusters provides
important information on structure formation process and on 
fundamental characteristics of our Universe. This field is at a 
crucial point of its development thanks to the very low sensitivity of present and future-generation 
instruments such as LOFAR\footnote{www.lofar.org}, \emph{e}ROSITA\footnote{www.mpe.mpg.de/eROSITA}, 
ASTRO-H\footnote{astro-h.isas.jaxa.jp} and CTA\footnote{www.cta-observatory.org}, to name but a few. 
Therefore, it is important to provide prospects for future surveys and
to understand in which way they can improve our current knowledge.
To this end, a complete cosmological mock sample of galaxy
clusters is needed. This can be realized either using analytical 
expressions for the cluster mass function (e.g., \citealp{2001MNRAS.321..372J}) or cosmological cluster 
simulations. In this work, we make use of the recent MultiDark $N$-body
simulation (\citealp{2011arXiv1104.5130P}; www.multidark.org).

@@

In next Section~\ref{sec:2}, we introduce the MultiDark simulation
along with our final cluster sample. We assign to each of our clusters  
the phenomenologically constructed gas density profile in 
Section~\ref{sec:3}. In Sections~\ref{sec:4} and 
\ref{sec:5}, we show that this approach can successfully reproduce 
the known X-ray and SZ cluster characteristics. We present the
resulting mock cluster catalogs in Section~\ref{sec:6}, and eventually 
summarize our conclusions in Section~\ref{sec:7}.
In this work, the cluster mass $M_{\Delta}$ and radius $R_{\Delta}$ are defined 
with respect to a density that is $\Delta=200$ or $\Delta=500$ times the \emph{critical} 
density of the Universe. We adopt density parameters of $\Omega_{\rmn{m}}=0.3$, 
$\Omega_{\Lambda}=0.7$ and today's Hubble constant of 
$H_0 = 100 \times h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.
